### 2024-12-04 manuscript.tex
Update on Overleaf.

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to a lack of surface and in situ observations and being a lower
priority region for numerical weather prediction (NWP) and climate model
development because of its distance from populated areas.  Nevertheless,
radiation biases and changes over an area of its size have a substantial
influence on the global climate, and the SO is an important part of the global
ocean conveyor belt.  Marine clouds have a disproportionate effect on top of
atmosphere (TOA) SW radiation due to the relatively low albedo of the sea
surface.  The relative longitudinal symmetry of the SO means that model cloud
biases tend to be similar across longitudes.  Here, we conventionally refer to
the SO as ocean regions south of 40°S, low-latitude SO as 40--55°S and
high-latitude SO as south of 55°S.

@@

In general, sea surface temperature (SST) biases in the SO can originate either
in the atmosphere, caused by too much shortwave heating of the surface, too
little longwave cooling of the surface, or in the ocean circulation.
Interactions of both are also possible, for example SST affecting clouds and
clouds affecting the surface radiation. \cite{zhang2023} have shown that SST
biases have improved in CMIP6 compared to CMIP5 [relative to the European
Centre for Medium-Range Weather Forecasts (ECMWF) Reanalysis 5; ERA5], with SST
overall increasing in the later CMIP phase. However, over the SO this resulted
in an even higher positive bias, especially in the Atlantic Ocean (AO) sector
of the SO, increasing by up to 1°C.  \cite{luo2023} identified that the SO SST
bias in an ensemble of 18 CMIP6 models originates not from the surface heat and
radiation fluxes (using reanalyses as a reference), but from a warm bias in the
Northern Atlantic Deep Water.

@@

The Vaisala CL51 and CT25K (photos in Fig. \ref{fig:map}d, e) are ceilometers
operating at a near-infrared wavelength of 910 nm and 905 nm, respectively.
The CL51 can also be configured to emulate the Vaisala CL31. The maximum range is
15.4 km (CL51), 7.7 km (CL31 emulation mode with 5 m vertical resolution), and
7.5 km (CT25K). The vertical resolution is 10 m (5 m configurable) in CL51 and
30 m in CT25K observations. The sampling (temporal) resolution is configurable, and in our
datasets is approximately 6 s for CL51 on AA15‐16, 16 s for CT25K on MARCUS and
MICRE, 36 s for CL51 on RV \emph{Polarstern}, and about 2.37 s for CL51 with
CL31 emulation on TAN1502. The wavelength of 910 nm is affected by water vapour
absorption of about 20\% in the mid-latitudes \citep{wiegner2015,wiegner2019},
but we do not expect this to be a significant issue as explained in
\cite{kuma2021}.  The instrument data files containing raw uncalibrated
backscatter were first converted to Network Common Data Form (NetCDF) with
cl2nc (\url{https://github.com/peterkuma/cl2nc}) and then processed with the
ALCF (Section \ref{sec:alcf}) to produce absolutely calibrated attenuated
volume backscattering coefficient (AVBC), cloud mask, cloud occurrence by
height, and the total cloud fraction. Because the CT25K uses a very similar
wavelength to CL51, equivalent calculations as for CL51 were done assuming a
wavelength of 910 nm. The Vaisala CL51 and CT25K instruments were used on most
of the voyages and stations analysed here. Fig.  \ref{fig:example}a shows an
example of AVBC derived from the CL51 instrument data.

@@

Here, we use a free-running (i.e., \emph{not} nudged or using prescribed SST)
coupled GSRM simulation made for the purpose of climate projection.  nextGEMS
has so far produced four cycles of model runs. We used a Cycle 3 run
\emph{ngc3028} produced in 2023 \citep{nextgems2023a,nextgems2023b} for a model
time period of 20 January 2020 to 22 July 2025, of which we analysed the period 2021--2024 (inclusive). While a Cycle 4 run was available, we could not
use it due to a lack of availability of the necessary variables. The horizontal
resolution of ngc3028 is about 5 km.  The model output is available on 90
vertical levels and 3-hourly instantaneous temporal resolution.  Unlike current
general circulation models (GCMs), the storm-resolving version of ICON does not
use convective and cloud parameterization but relies on explicit simulation of
convection and clouds on the model grid. While this makes the code development
simpler without having to rely on uncertain parameterizations, it can miss
smaller-scale clouds below the grid resolution.  Turbulence and cloud
microphysics are still parameterized in this model.

@@

The Modern-Era Retrospective analysis for Research and Applications, Version 2
(MERRA-2) is a reanalysis produced by the Global Modeling and Assimilation
Office at the NASA Goddard Space Flight Center \citep{gelaro2017}.  It uses
version 5.12.4 of the Goddard Earth Observing System (GEOS) atmospheric model
\citep{rienecker2008,molod2015}. The reanalysis output analysed here is
available at a spatial resolution of 0.5° of latitude and 0.625° of longitude,
which is about 56 km in the North--South direction and 35 km in the East--West
direction at 60°S. The number of vertical model levels is 72. Here, we use the
following products: 1-hourly instantaneous 2D single-level diagnostics
(M2I1NXASM) for 2-m temperature and humidity; 3-hourly instantaneous 3D
assimilated meteorological fields (M2I3NVASM) for cloud quantities, pressure,
and temperature; 1-hourly average 2D surface flux diagnostics (M2T1NXFLX) for
precipitation; and 1-hourly average 2D radiation diagnostics
(M2T1NXRAD) for radiation quantities \citep{merra2}.

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, ERA5 and MERRA-2 (Fig.
\ref{fig:cloud-occurrence-panel}). In addition, we aggregated the data sources
(voyages and stations) by calculating the averages and percentiles of all
individual profiles, presented in Fig. \ref{fig:cloud-occurrence}. The analysis
shows that the total cloud fraction (determined as the fraction of profiles
with clouds at any height in the lidar cloud mask) is underestimated in ICON
by about 10\% and in the reanalyses by about 20\%. When analysed by
height, ICON overestimates cloud occurrence below 1 km and underestimates it
above, MERRA-2 underestimates cloud occurrence at all heights, especially near
the surface, and ERA5 simulates cloud occurrence relatively well above 1 km,
but strongly underestimates it near the surface.  We note that fog or
near-surface clouds are strongly underestimated in the reanalyses (fog and clouds are
both included in the cloud occurrence).  As shown in Fig.
\ref{fig:cloud-occurrence-panel}, the biases are relatively consistent across
the data sources and longitudes. We conclude that the ICON results ch the observations better than the reanalyses in this metric.

>>>>

The Southern Ocean (SO) is known to be a problematic region for climate model
biases due to a lack of surface and in situ observations and being a lower
priority region for numerical weather prediction (NWP) and climate model
development because of its distance from populated areas.  Nevertheless,
radiation biases and changes over an area of its size have a substantial
influence on the global climate, and the SO is an important part of the global
ocean conveyor belt. In general, marine clouds have a disproportionate effect on top of
atmosphere (TOA) SW radiation due to the relatively low albedo of the sea
surface.  The relative longitudinal symmetry of the SO means that model cloud
biases tend to be similar across longitudes.  Here, we conventionally refer to
the SO as ocean regions south of 40°S, low-latitude SO as 40--55°S and
high-latitude SO as south of 55°S.

@@

In general, sea surface temperature (SST) biases in the SO can originate either
in the atmosphere, caused by too much shortwave heating of the surface, too
little longwave cooling of the surface, or in the ocean circulation.
Interactions of both are also possible, for example SST affecting clouds and
clouds affecting the surface radiation. Using the European
Centre for Medium-Range Weather Forecasts (ECMWF) Reanalysis 5 (ERA5) as a reference, \cite{zhang2023} have shown that SST
biases have improved in CMIP6 compared to CMIP5, with SST
overall increasing in the later CMIP phase. However, over the SO this resulted
in an even higher positive bias, especially in the Atlantic Ocean (AO) sector
of the SO, increasing by up to 1°C.  \cite{luo2023} identified that the SO SST
bias in an ensemble of 18 CMIP6 models originates not from the surface heat and
radiation fluxes (using reanalyses as a reference), but from a warm bias in the
Northern Atlantic Deep Water.

@@

The Vaisala CL51 and CT25K (photos in Fig. \ref{fig:map}d, e) are ceilometers
operating at a near-infrared wavelength of 910 nm and 905 nm, respectively.
The CL51 can also be configured to emulate the Vaisala CL31. The maximum range is
15.4 km (CL51), 7.7 km (CL31 emulation mode with 5 m vertical resolution), and
7.5 km (CT25K). The vertical resolution is 10 m (5 m configurable) in CL51 and
30 m in CT25K observations. The sampling (temporal) resolution is configurable, and in our
datasets is approximately 6 s for CL51 on AA15‐16, 16 s for CT25K on MARCUS and
MICRE, 36 s for CL51 on RV \emph{Polarstern}, and about 2.37 s for CL51 with
CL31 emulation on TAN1502. The wavelengths of 905 and 910 nm are affected by water vapour
absorption of about 20\% in the mid-latitudes \citep{wiegner2015,wiegner2019}, with 910 nm affected more strongly,
but we do not expect this to be a significant issue as explained in
\cite{kuma2021}.  The instrument data files containing raw uncalibrated
backscatter were first converted to Network Common Data Form (NetCDF) with
cl2nc (\url{https://github.com/peterkuma/cl2nc}) and then processed with the
ALCF (Section \ref{sec:alcf}) to produce absolutely calibrated attenuated
volume backscattering coefficient (AVBC), cloud mask, cloud occurrence by
height, and the total cloud fraction. Because the CT25K uses a very similar
wavelength to CL51, equivalent calculations as for CL51 were done assuming a
wavelength of 910 nm. The Vaisala CL51 and CT25K instruments were used on most
of the voyages and stations analysed here. Fig.  \ref{fig:example}a shows an
example of AVBC derived from the CL51 instrument data.

@@

Here, we use a free-running (i.e., the weather situation in the model does not correspond to reality)
coupled GSRM simulation made for the purpose of climate projection.  nextGEMS
has so far produced four cycles of model runs. We used a Cycle 3 run
\emph{ngc3028} produced in 2023 \citep{nextgems2023a,nextgems2023b} for a model
time period of 20 January 2020 to 22 July 2025, of which we analysed the period 2021--2024 (inclusive). While a Cycle 4 run was available, we could not
use it due to a lack of availability of the necessary variables. The horizontal
resolution of ngc3028 is about 5 km.  The model output is available on 90
vertical levels and 3-hourly instantaneous temporal resolution.  Unlike current
general circulation models (GCMs), the storm-resolving version of ICON does not
use convective and cloud parameterization but relies on explicit simulation of
convection and clouds on the model grid. While this makes the code development
simpler without having to rely on uncertain parameterizations, it can miss
smaller-scale clouds below the grid resolution.  Turbulence and cloud
microphysics are still parameterized in this model.

@@

The Modern-Era Retrospective analysis for Research and Applications, Version 2
(MERRA-2) is a reanalysis produced by the Global Modeling and Assimilation
Office at the NASA Goddard Space Flight Center \citep{gelaro2017}.  It uses
version 5.12.4 of the Goddard Earth Observing System (GEOS) atmospheric model
\citep{rienecker2008,molod2015}. Non-convective clouds (condensation, autoconversion, and evaporation) are parametrised using a prognostic scheme \citep{bacmeister2006}, and sub-grid cloud fraction is determined using total water distribution and a critical relative humidity threshold. The reanalysis output analysed here is
available at a spatial resolution of 0.5° of latitude and 0.625° of longitude,
which is about 56 km in the North--South direction and 35 km in the East--West
direction at 60°S. The number of vertical model levels is 72. Here, we use the
following products: 1-hourly instantaneous 2D single-level diagnostics
(M2I1NXASM) for 2-m temperature and humidity; 3-hourly instantaneous 3D
assimilated meteorological fields (M2I3NVASM) for cloud quantities, pressure,
and temperature; 1-hourly average 2D surface flux diagnostics (M2T1NXFLX) for
precipitation; and 1-hourly average 2D radiation diagnostics
(M2T1NXRAD) for radiation quantities \citep{merra2}.

@@

We used the ALCF to derive cloud occurrence by height and the total cloud
fraction from observations, ICON, ERA5 and MERRA-2 (Fig.
\ref{fig:cloud-occurrence-panel}). In addition, we aggregated the data sources
(voyages and stations) by calculating the averages and percentiles of all
individual profiles, presented in Fig. \ref{fig:cloud-occurrence}. The analysis
shows that the total cloud fraction (determined as the fraction of profiles
with clouds at any height in the lidar cloud mask) is underestimated in ICON
by about 10\% and in the reanalyses by about 20\%. When analysed by
height, ICON overestimates cloud occurrence below 1 km and underestimates it
above, MERRA-2 underestimates cloud occurrence at all heights, especially near
the surface, and ERA5 simulates cloud occurrence relatively well above 1 km,
but strongly underestimates it near the surface.  We note that fog or
near-surface clouds are strongly underestimated in the reanalyses (fog and clouds are
both included in the cloud occurrence).  As shown in Fig.
\ref{fig:cloud-occurrence-panel}, the biases are relatively consistent across
the data sources and longitudes. We conclude that the ICON results  the observations better than the reanalyses in this metric.
